\documentclass[12pt]{article}

%Input preamble, math commands, environments, etc. from a saingle file. 
\input{HelperFiles/BenPorath}
\begin{document}

\title{The Accumulation of Human Capital over the Life Cycle: Lessons from the Yoram Ben-Porath Model}
\author{Prepared by: Jorge L. Garc\'{i}a \\ The University of Chicago}
\date{This Draft: \today}
\maketitle

\begin{flushright}
\textbf{Cite of Mincer, 1958.}
\end{flushright}


\begin{abstract}
\noindent For several years, economists have worried about theories of income and its prediction of observable outcomes or constructs of social interest. Particular interest they have paid to theories relating the distribution of income and the distribution of abilities (\citet{staehle1943ability} offers the first formal treatment on this topic in Economics). The most relevant ingredient of this analysis is human capital and its evolution. \citet{mincer1958investment} and \citet{becker1962investment} are the two main predecessors of the analysis of human capital investment. The former firstly asked why the distributions of abilities and income differed and argued for market compensation for different worker traits as principal causes. The latter established the first self-contained theoretical analysis of human capital investment. \citet{ben1967production} generalized \citet{becker1962investment} to a dynamic context that allows for a very rich analysis of the production of human capital. Specifically, it enables to analyze variations in (i) specialization periods; (ii) production functions of human capital; (iii) time horizons; (iv) rates of return. Based on the model in \citet{ben1967production}, we provide a framework to analyze the evolution of earnings and human capital under a wide variety of scenarios. We hope to guide researchers on the consequences of their human capital production and accumulation modeling choices.
\end{abstract}

\input{Sections/BasicModel}

\clearpage
\bibliographystyle{chicago}
\bibliography{BibtexFiles/BenPorath}



\end{document}
